---
title: "Preliminary_results.Rmd"
author: "Camille DELABAERE"
date: "2024-12-01"
output: 
  html_document
---

### 1/ Title : Preliminary_results of [Cooperation between arbuscular mycorrhizal fungi and plant growth-promoting bacteria and their effects on plant growth and soil quality](https://peerj.com/articles/13080/)

### 2/ Group members : Keve Toth ; Jakov Cancar ; Camille Delabaere

### 3/ Introduction. Why your project is interesting? & What is known already about it?

We chose this scientific paper because we were interested in the knowledge we could gain about the relationship between plants & microorganisms *(here arbuscular mycorrhizal fungi (AMF) & plant growth-promoting bacteria (PGPB))*. A way to understand how the growth of plants & physico chemical properties of the soil are influenced by different treatments. Finaly, as "Almost 25% of the Earth’s biodiversity is composed of soil microorganisms (...)" (Fierer, 2017; Wagg et al., 2019; Whitman, Coleman & Wiebe, 1998), we thought we would acquire knowledge that we can later on, apply during our career.

What we know already :

**AMF** - present to 70% of the terrestrial plants *"Arbuscular mycorrhizal fungi (AMF), (...) maintain symbiotic relationships with over 70% of terrestrial plants and provide nutrients and water to plants in exchange for sugars through their arbuscules (...)"* (Schussler, Schwarzott & Walker, 2001; Wagg et al., 2019) and favorise the exchange of nutrients/sugars thanks to arbuscules. It enhances resilience of plants for drought, heavy metals & pathogens.

**PGPR** - include B (Bacillus spp…), contribute to the fixation of N, solubilisation of P, production of phytohormones & systemic resistance. *Bacillus spp. promotes plant growth by fixing N, solubilizing and mineralizing P and other nutrients, stimulating phytohormones, producing siderophores, inducing systemic resistance (ISR), and enhancing their tolerance to abiotic stresses* (Bonfante & Genre, 2010; Saxena et al., 2020). AMF & PGPR provide ecological solutions to improve health soil & plants productivity.

### 4/ Objectives and hypotheses

The <u>objectives</u> are to analyse if arbuscular mycorrhizal fungi\* AMF\* & plant growth-promoting rhizobacteria *PGRB* have any synergy effects on nutrients uptakes or plant growth to a more effective extent than for each microorganism studied alone.

The <u>hypotheses</u> are 1) AMF and PGPR could mutually symbiose and enhance the plant growth of Elymus nutans Griseb, & (2) the co-existence of AMF and PGPR could improve plant traits and soil quality better than their individual applications.

........................................

### 5. Methods.

### 5.a Origin & structure of the data.

It comes from the scientific article [Cooperation between arbuscular mycorrhizal fungi and plant growth-promoting bacteria and their effects on plant growth and soil quality](https://peerj.com/articles/13080/) by Yu L, Zhang H, Zhang W, Liu K, Liu M, Shao X. 2022.

As we wanted to make the data more readable, it was sorted by plant_param and soil_param.

Loading of Data
```{r}
Data_Soil_param <- read.csv(file.choose())
Data_Plant_param <- read.csv(file.choose())
```

Displays tables in the document
```{r}
knitr::kable(Data_Soil_param, format = "html") 
knitr::kable(Data_Plant_param, format = "html")
```

### 5.b What have you done? And what did you use to do it? Step by step, so anyone could do it again by reading this part.

### 6. Results. Figures and tables with captions and description of what do they mean. Never include citations in this part. This is only your work.

### 7. Discussion. From your objectives.

8.  Conclusions.

.........................................
